---
title: "Location of the Lie EDA"
author: "Sofia Bliss-Carrascosa"
date: "2/28/2022"
output: pdf_document
---

```{r loadpackages}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

```{r}
loc_tags <- read.csv("first700loctags.csv") %>%
  filter(location != "")
```

```{r}
loc_tags %>%
  group_by(location) %>%
  summarize(count = n()) %>%
  kable()
```
Proposed categories:
- Debate (campaign debates, panel discussion)
- Statement on TV (any televised interview, video remarks)
- Interview (radio interview and conversation with reporters)
- Public interaction (town hall, forum)
- Advertisement (campaign, television, radio, and unspecified)
- Social Media (non-campaign video: tweets, facebook posts, etc)
- Written communication (own website, emails, fliers, own writing/blog posts, letters, mailers, press releases)
- Press release (press briefing)

```{r}
proposed_loc_tags <- loc_tags %>%
  mutate(prop_location = case_when(location == "panel discussion" ~ "debate",
                              location == "debate" ~ "debate",
                              location == "video remarks" ~ "statement on tv",
                              location == "statement on tv" ~ "statement on tv",
                              location == "televised address" ~ "statement on tv",
                              location == "radio interview" ~ "interview",
                              location == "interview" ~ "interview",
                              location == "town hall" ~ "public interaction",
                              location == "forum" ~ "public interaction",
                              location == "advertisement" ~ "advertisement",
                              location == "attack ad" ~ "advertisement",
                              location == "online advertisement" ~ "advertisement",
                              location == "radio advertisement" ~ "advertisement",
                              location == "tv advertisement" ~ "advertisement",
                              location == "own website" ~ "written communication",
                              location == "email" ~ "written communication",
                              location == "email newsletter" ~ "written communication",
                              location == "own writing" ~ "written communication",
                              location == "letter" ~ "written communication",
                              location == "mailer" ~ "written communication",
                              location == "press release" ~ "written communication",
                              location == "email // own writing" ~ "written communication",))

proposed_loc_tags %>%
  group_by(prop_location) %>%
  summarize(count = n()) %>%
  kable()
```


